### 2022-02-12 symposium.tex
Type-Safe Regular Expressions: related work

The second implementation of our runtime component takes a completely different approach.
Instead of analyzing regular expressions at runtime, to reconstitute type-level analysis results, we use implicit resolution to \emph{synthesize} a type-specialized program.
That synthesis is type-directed: it generates a single-purpose program that sanitized the output of the regex engine for a particular regular expression.

Our implementation consists of a type class, |Sanitizer|, and three implicit definitions, which will guide the compiler into generating the correct |Sanitizer[T]|, for any tuple |T| with |String| and |Option[String]| elements.
That design is analogous to the recursive implicit
\Cref{sec:recursive-implicit-resolution}

The |Sanitizer| type class defines a single method that mutates an array in place to warp its nullable elements into options:

@@

From a design standpoint, this solution is appealing as it leads to no duplicated code and no duplicated computations: we perform our capture group analysis at the type-level, once and for all, and synthesize code accordingly.
However, this solution comes with an additional cost in terms of compilation time.
In the following section, we evaluate the performance of our library and discuss the trade-offs between the first and second implementation of our runtime component.

@@

@@

They did it all: \citep{spishak2012type}
Also this talk, from Stephanie: \citep{weirich2014examples}

>>>>

The second implementation of our runtime component takes a completely different approach.
Instead of analyzing regular expressions at runtime, to reconstitute type-level analysis results, we use implicit resolution to \emph{synthesize} a type-specialized runtime.
That synthesis is type-directed: it generates a single-purpose program that sanitized the output of the regex engine for a particular regular expression.

Our implementation consists of a type class, |Sanitizer|, and three implicit definitions, which will guide the compiler into generating the correct |Sanitizer[T]|, for any tuple |T| with |String| and |Option[String]| elements.
The |Sanitizer| type class defines a single method that mutates an array in place to warp its nullable elements in options:

@@

From a design standpoint, this solution is appealing as it leads to no duplicated code and no duplicated computations: we perform our capture group analysis at the type-level, once and for all, and synthesize code accordingly.
This combination of type-level programming and implicit-based code synthesis is reminiscent of the compile-time variant of multi-stage metaprogramming \citep{stucki2018practical}.
In those terms, our approach corresponds to a multi-staged program whose stage 0 happens entirely within Scala's type checker.

In the following section, we evaluate the performance of our library and discuss the trade-offs between the first and second implementation of our runtime component.

@@

XXX

@@

\citeauthor{spishak2012type} developed a type system to validate regular expression syntax and capturing group extraction in Java programs \citep{spishak2012type}.
Their tool takes the form of a compiler plugin that enhances Java's type system to track the number of capturing groups using type annotations.
Unlike our system, their are not concerned with null safety and does not detect optional capturing groups.
In practice, their tool operates similarly to a linter: it reports false positives and sometimes requires manual annotations.

\citep{mainland2007s}'s quasiquote

Also this talk, from Stephanie: \citep{weirich2014examples}
dependently-typed flavor of GHC Haskell which requires no less than 20 language extensions, whereas our implementation compiles with vanilla Scala~3.
